\huge{Comment on `The Limits of Cosmic Shear'}

### 2016-12-09 cfhtlens.tex
Added MK's final comments

We note that \citet{kitching/etal:2016} use their `Flat' approximation for their `Full' non-Limber analysis of the CFHTLenS data.   This is because on a sphere, correlation functions cannot be related to the power spectrum by a Hankel transform (e.g. equation~\ref{eqn:xiGG}) as the power spectrum is not defined for non-integer $\ell$.   In Figure~\ref{fig:Cl_xi}, however, we show that this chosen version of a flat-sky approximation serves to exacerbate the tension in theoretical models when a flat-sky solution is compared with the spherical-sky solution in the Limber approximated case.  This is not the case when considering the standard $T_\ell = 1$ flat-sky approximation that has been used in the majority of cosmic shear studies to date.  We would therefore argue that it is more appropriate to use a $T_\ell = 1$ flat-sky approximation when carrying out a `Full' non-Limber analysis of the shear correlation function.

>>>>

We note that \citet{kitching/etal:2016} use their `Flat' approximation for their `Full' non-Limber analysis of the CFHTLenS data.   This is because on a sphere, correlation functions cannot be related to the power spectrum by a Hankel transform (e.g. equation~\ref{eqn:xiGG}) as the power spectrum is not defined for non-integer $\ell$.   In Figure~\ref{fig:Cl_xi}, however, we show that this chosen version of a flat-sky approximation serves to exacerbate the tension in theoretical models when a flat-sky solution is compared with the spherical-sky solution in the Limber approximated case.  This is not the case when considering the standard $T_\ell = 1$ flat-sky approximation that has been used in the majority of cosmic shear studies to date.  We would therefore argue that it is more appropriate to use a $T_\ell = 1$ flat-sky approximation when carrying out a `Full' non-Limber analysis of the shear correlation function obtained from the Hankel transformation (equation~\ref{eqn:xiGG}).

### 2016-12-09 theory.tex
Added MK's final comments

 \subsection{Cosmic shear without the Limber Approximation}
In collecting our comments on \citet{kitching/etal:2016} we contacted colleagues who had previously tested the impact of the Limber approximation on cosmic shear studies by comparing a Limber-approximated cosmic shear power spectrum with an exact calculation.  The majority of these analyses remained unpublished in peer reviewed journals as no significant deviation was found\footnote{See for example appendix M of the 2010 PhD thesis from Donghui Jeong (\url{http://www.personal.psu.edu/duj13/dissertation/djeong_diss_appM.pdf}) and Chapter 7 of 2012 PhD thesis from Nicolas Van de Rijt (\url{http://ipht.cea.fr/Docspht/articles/t12/080/public/these_vanderijt2012.pdf}).}.  One notable exception, however, is \citet{giannantonio/etal:2012} who find that for a cosmic shear power spectrum, the extended Limber approximation is consistent with the exact calculation for $\ell>5$, with any differences well within cosmic-variance errors for a Euclid-like survey\footnote{\citet{kitching/etal:2016} comment on this result, but indicate that it is in error owing to the application of a fixed low-$\ell$ limit.  \citet{giannantonio/etal:2012} however state that this limit is only applied in a positive curvature case.}.  They conclude that the Limber approximation works well for cosmic shear owing to the wide redshift distribution of the lensed sources.  This is in contrast to the case of galaxy power spectra,  where narrower redshift bins result in a significant difference between the Limber and non-Limber cases \citep[see also][]{simon/2007}.   Given this body of work, we argue that the assertion by \citet{kitching/etal:2016} that the majority of cosmic shear data analyses to date have made an `axiomatic assumption (i.e. an unquestioned and untested assumption at the beginning of the analysis)'  is unwarranted.  Not only is there published work that has questioned and tested the Limber approximation for the cosmic shear power spectrum, individual teams have also verified this result internally for their own cosmic shear analyses.

Given the insignificant effect of using the Limber approximation found by various groups, the ability to calculate an exact non-Limber solution has unfortunately not been maintained in many theoretical codes over the years.  Instead these codes have evolved to account for critical approximations such as the impact of baryon feedback on the matter power spectrum and non-zero neutrino masses \citep[see for example][]{joudaki/etal:2016, mead/etal:2016}.    Two exceptions to this are recent updates to CLASS \citep{blas/lesgourgues/tram:2011,audren/etal:2013} and CosmicFish \citep{raveri/etal:2016}.  On-going code comparisons have, however, revealed differences between the calculated theoretical models that are larger than the expected impact of Limber or non-Limber approximation.   We are therefore unable to readily determine non-Limber theoretical calculations for the shear-correlation function to directly compare with the results of \citet{kitching/etal:2016}.   Instead we refer the reader to Figure 3 of \citet{giannantonio/etal:2012} to draw their own conclusions about the impact of this effect.

>>>>

 \subsection{Cosmic shear without the Limber Approximation}
In collecting our comments on \citet{kitching/etal:2016} we contacted colleagues who had previously tested the impact of the Limber approximation on cosmic shear studies by comparing a Limber-approximated cosmic shear power spectrum with an exact calculation.  The majority of these analyses were not published in peer reviewed journals as no significant deviation was found\footnote{See for example appendix M of the 2010 PhD thesis from Donghui Jeong (\url{http://www.personal.psu.edu/duj13/dissertation/djeong_diss_appM.pdf}).}.  Two notable exceptions, however, are \citet{giannantonio/etal:2012} and \citet{bernardeau/etal:2012}.  Figure 1 of \citet{bernardeau/etal:2012} shows that the Baseline Limber approximated cosmic shear power spectrum is accurate to better than 1\% at $\ell > 8$.  \citet{giannantonio/etal:2012} also find that the extended Limber approximation is consistent with the exact calculation for their minimum tested $\ell>5$, with any differences well within cosmic-variance errors for a Euclid-like survey\footnote{\citet{kitching/etal:2016} comment on this result, but indicate that it is in error owing to the application of a fixed low-$\ell$ limit.  \citet{giannantonio/etal:2012} however state that this limit is only applied in a positive curvature case.}.  They conclude that the Limber approximation works well for cosmic shear owing to the wide redshift distribution of the lensed sources.  This is in contrast to the case of galaxy power spectra,  where narrower redshift bins result in a significant difference between the Limber and non-Limber cases \citep[see also][]{simon/2007}.

Given this body of work, we argue that the assertion by \citet{kitching/etal:2016} that the majority of cosmic shear data analyses to date have made an `axiomatic assumption (i.e. an unquestioned and untested assumption at the beginning of the analysis)'  is unwarranted.  Not only is there published work that has questioned and tested the Limber approximation for the cosmic shear power spectrum, individual teams have also verified this result internally for their own cosmic shear analyses.

Given the insignificant effect of using the Limber approximation found by various groups, the ability to calculate an exact non-Limber solution has unfortunately not been maintained in many theoretical codes over the years.  Instead these codes have evolved to account for critical approximations such as the impact of baryon feedback on the matter power spectrum and non-zero neutrino masses \citep[see for example][]{joudaki/etal:2016, mead/etal:2016}.    Two exceptions to this are recent updates to CLASS \citep{blas/lesgourgues/tram:2011,audren/etal:2013} and CosmicFish \citep{raveri/etal:2016}.  On-going code comparisons have, however, revealed differences between the calculated theoretical models that are larger than the expected impact of Limber or non-Limber approximation.   We are therefore unable to readily determine non-Limber theoretical calculations for the shear-correlation function to directly compare with the results of \citet{kitching/etal:2016}.   Instead we refer the reader to Figure 3 of \citet{giannantonio/etal:2012} and Figure 1 of  \citet{bernardeau/etal:2012} to draw their own conclusions about the impact of this effect.
